---
title: "Reproducible research - Week 2 Assignment"
author: "François de BELLEFON"
date: "21 mai 2017"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
library("dplyr")
library("ggplot2")
library("gridExtra")
library("lattice")
opts_knit$set(root.dir = "C:/Coursera/Reproduceable/Week2/repdata_data_activity")
Sys.setenv(LANG = "en")
Sys.setlocale("LC_ALL","English")
```

#Loading and processing the data

We're reading the data from the csv file. It has a header and its NA values are "NA".
```{r LoadingData}
act <- read.csv("activity.csv",header=TRUE,na.strings = "NA")
```

#What is mean total number of steps taken per day?

##Steps taken per day

Let's calculate the total number of steps taken per day :

```{r TotalSteps}
act_date <- act %>% group_by(date)

date_sum <- act_date %>% summarise(stepssum = sum(steps,na.rm=TRUE)) 


date_sum
```
##Histogram
Let's draw the histogram of it :
```{r TotalStepsHistogram}
stepsHist <- ggplot(data=date_sum, aes(stepssum)) + geom_histogram() + xlab("Number of steps") + ggtitle("Histogram of the total number of steps taken each day")
stepsHist
```

##Mean per day
Let's calculate the mean of the total number of steps taken per day :

```{r MeanTotalSteps}
date_mean <- act_date %>% summarise(stepsmean = mean(steps,na.rm=TRUE)) 
date_mean
```

##Median per day
Let's calculate the median of the total number of steps taken per day :
```{r MedianTotalSteps}
date_median <- act_date %>% summarise(stepsmed = mean(steps,na.rm=TRUE)) 
date_median
```

#Average daily activity pattern
##Time plot series
Let's plot a time series of the 5-minutes interval and the average number of steps taken, averaged across all days :
```{r DailyActivityPatternPlot}
act_int <- act %>% group_by(interval)
int_mean <- act_int %>% summarise(stepintmean = mean(steps, na.rm=TRUE))
plot(int_mean,type="l")
```
##Maximum number of steps per interval
The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps is the following :

```{r Max5MinInterval}
int_mean[which.max(int_mean$stepintmean),]$interval
```

#Imputing missing values

##Total number of missing values
Here is the total number of missing values in the dataset :
```{r TotalMissingValues}
sum(is.na(act$steps))
```

##Strategy to replace the missing values :
For each missing values, replace it by the mean for the corresponding 5 minutes interval that we calculated earlier.


##Implementing this strategy
```{r ReplacingMissingValues}
# Initialize this new dataset :
actWithoutNA <- act


for(i in 1:length(actWithoutNA$steps)){
  if(is.na(actWithoutNA$steps[i])){
    actWithoutNA$steps[i] <- int_mean$stepintmean[int_mean$interval == actWithoutNA$interval[i]]
  }
}
actWithoutNA
```

##Comparing results

###Histogram
Let's draw an histogram of the total number of steps per day.

For comparison purposes, we add the old histogram (with the NA values) next to it.
```{r TotalNumberHistogram}
act_nona_date <- actWithoutNA %>% group_by(date)

date_nonna_sum <- act_nona_date %>% summarise(stepsnonnasum = sum(steps,na.rm=TRUE)) 

stepsHistWithoutNA <- ggplot(data=date_nonna_sum, aes(stepsnonnasum)) + geom_histogram() + xlab("Number of steps") + ggtitle("Histogram of the total number of steps taken each day", subtitle = "NA replaced")

grid.arrange(stepsHist + ylim(0,12.5),stepsHistWithoutNA)
```


###Mean of values
Let's calculate the mean of values with replaced NAs :
```{r MeanWithoutNA}
date_nonna_mean <- act_nona_date %>% summarise(stepsmean = mean(steps,na.rm=TRUE)) 
date_nonna_mean
```


```{r MeanComparison}
MergedMean <- merge(date_nonna_mean,date_mean,by="date")
```




### Median of the values :
Let's calculate the median of values with replaced NAs :
```{r MedianWithoutNA}
date_nonna_median <- act_nona_date %>% summarise(stepsmed = median(steps,na.rm=TRUE))

```




As we can see, these values differ from the one calculated precedently.



#Differences in activity patterns between weekdays and weekends


Let's create a dataset indicating if a date is a weekday or weekend :
```{r WeekDayWeekEnd}
actWithoutNAwkd <- actWithoutNA %>% mutate(wkd = weekdays(as.Date(date)))
actWithoutNAwkd$wkd[actWithoutNAwkd$wkd %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")] <- "weekday"
actWithoutNAwkd$wkd[actWithoutNAwkd$wkd %in% c("Saturday","Sunday")] <- "weekend"
actWithoutNAwkd
```

##Average number of steps taken per day, averaged across all week days and week end days
```{r ComparisonWeekDaysWeekEnd}
actWithoutNAwkdgrp <- actWithoutNAwkd %>% group_by(wkd,interval) %>% summarise(stepswkdint = sum(steps,na.rm=TRUE))
actWithoutNAwkdgrp
xyplot(stepswkdint~interval|wkd, data = actWithoutNAwkdgrp, type = "l")
```